Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-22 robdml.tex
tex

Given a set of data points, a metric gives a distance value between each pair
of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
As an extensively studied task~\cite{revisiting}, DML has a wide range of
applications such as image retrieval~\cite{imagesim2} and face
recognition~\cite{facenet,domainface}, as well as a wide influence on
the core idea of other areas such as self-supervised learning~\cite{dmlreality}.

Despite the advancements in this field thanks to deep neural
networks, recent studies find DML models vulnerable to
adversarial attacks, where an imperceptible perturbation can incur unexpected
retrieval, or covertly manipulate the results~\cite{advrank,advorder}.
%
Such vulnerability raises security, safety, and fairness concerns in the
DML applications.
%
For example, impersonation or recognition evation are possible for a
vulnerable DML-based face-identification system.
%
To counter the attacks (\ie, mitigating the vulnerability), the
adversarial robustness of DML models has to be improved via defense.

@@

% existing methods & problem
Several adversarial training-based defense methods for DML are proposed in the
literature~\cite{advrank,robrank}, which are inspired by the Madry's min-max
adversarial training~\cite{madry} because it is one of the most effective
defense methods for deep neural network classifiers.
%
However, it is widely known that such methods always suffer from greatly increased
training cost and a notable penalty in the model's retrieval performance on
benign (\ie, unperturbed) examples.
%
Meanwhile, the level of robustness that the previous methods have achieved
is still insufficient.
%
(1) the robustness level achieved by the existing methods is still insufficient
to counter the attacks;
%
(2) the direct adaptation of Madry's min-max adversarial training~\cite{madry}
will easily lead to model collapse due to producing very hard sample triplets;
%
(3) the adversarial training procedure is very time-consuming compared to the
training process of a regular DML model, but existing defense methods are
incompatible to acceleration methods like Free Adversarial
Training~\cite{freeat}.  \oo{rob, time, r1}.

>>>>

Given a set of data points, a \emph{metric} gives a distance value between each
pair of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
As an extensively studied task~\cite{revisiting}, DML has a wide range of
applications such as image retrieval~\cite{imagesim2} and face
recognition~\cite{facenet,domainface}, as well as a wide influence on the core
idea of other areas such as self-supervised learning~\cite{dmlreality}.

Despite the advancements in this field thanks to deep neural
networks, recent studies find DML models vulnerable to
adversarial attacks, where an imperceptible perturbation can incur unexpected
retrieval, or covertly manipulate the results~\cite{advrank,advorder}.
%
Such vulnerability raises security, safety, and fairness concerns in the
DML applications.
%
For example, impersonation or recognition evation are possible for a
vulnerable DML-based face-identification system.
%
To counter the attacks (\ie, mitigating the vulnerability), the
\emph{adversarial robustness} of DML models has to be improved via defense.

@@

% existing methods & problem
Existing defense methods for DML~\cite{advrank,robrank} are adversarial
training-based, inspired by Madry's \emph{min-max} adversarial
training~\cite{madry} because it is one of the most effective methods for deep
neural network classifiers.
%
Specifically, Madry's method involves a inner problem to \emph{maximize} the
loss by perturbing the inputs into adversarial examples, and an outer problem
to \emph{minimize} the loss by updating the model parameters.
%
However, in order to avoid model collapse, the existing DML defenses refrain
from directly adopting such paradigm, but instead use some indirect and
relatively weak objectives to merely increase the loss value to a certain
level.
%
As a result, the model is not learning from a strong adversary, and suffers
from low efficiency in gaining robustness within limited time. 
%
Furthermore, advertsarial training already suffers from
significant training time cost (compared to the normal training process).
%
Hence, time efficiency in gaining adversarial robustness is an inevitable
major issue.

We conjecture there is a certain intermediate level of hardness.

notable penalty in the model performance on benign (\ie, unperturbed) examples.

%
\oo{rev}
%
Meanwhile, the level of robustness that the previous methods have achieved
is still insufficient.
%
(1) the robustness level achieved by the existing methods is still insufficient
to counter the attacks;
%
(2) the direct adaptation of Madry's min-max adversarial training~\cite{madry}
will easily lead to model collapse due to producing very hard sample triplets;
%
(3) the adversarial training procedure is very time-consuming compared to the
training process of a regular DML model, but existing defense methods are
incompatible to acceleration methods like Free Adversarial
Training~\cite{freeat}.  \oo{rob, time, r1}.
